---
title: "Investigate the exponential distribution in R and compare it with the Central Limit Theorem"
author: "Glenn Kerbein"
date: "April 1, 2017"
output: pdf_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
The purpose of this analysis is to investigate the exponential distribution and compare it to the Central Limit Theorem. For this analysis, set ``lambda`` equal to ``0.2`` for all simulations. The investigation will compare the distribution of averages of 40 exponentials over 1000 simulations.

## Simulations

Set the simulation variables lambda, exponentials, and seed.

```{r echo=TRUE}
ECHO=TRUE
set.seed(9001)
lambda <- 0.2
exponentials <- 40
```

Run the simulations with variables.

```{r echo=TRUE}
simMeans = NULL
for (i in 1 : 1000) {
	simMeans <- c(
		simMeans,
		mean(
			rexp(
				exponentials,
				lambda
			)
		)
	)
}
```

## Sample Mean versus Theoretical Mean

### Sample Mean

Calculating the mean from the simulations gives the sample mean.

```{r echo=TRUE}
mean(simMeans)
```

### Theoretical Mean

The theoretical mean of an exponential distribution is ``lambda^(-1)``.

```{r echo=TRUE}
lambda ^ (-1)
```

### Comparison

```{r echo=TRUE}
abs(mean(simMeans)-lambda^(-1))
```
There is only a slight difference between the simulations sample mean and the exponential distribution theoretical mean.

## Sample Variance versus Theoretical Variance

### Sample Variance

Calculating the variance from the simulation means gives the sample variance.

```{r echo=TRUE}
var(simMeans)
```

### Theoretical Variance

The theoretical variance of an exponential distribution is `` ( lambda * sqrt(n) ) ^ (-2) ``.

```{r echo=TRUE}
( lambda * sqrt(exponentials) ) ^ (-2)
```

### Comparison

There is only a small difference between the simulations sample variance and the exponential distribution theoretical variance.

```{r echo=TRUE}
abs(var(simMeans)-(lambda * sqrt(exponentials))^(-2))
```

## Distribution

Below is a density histogram of the 1000 simulations. There is an overlay with a normal distribution that has a mean of ``lambda ^ (-1)`` and standard deviation of ``( lambda * sqrt(n) ) ^ (-1)``, the theoretical normal distribution for the simulations.

```{r echo=TRUE}
library(ggplot2)
ggplot(
	data.frame(y=simMeans),
	aes(x=y)
)
	+ geom_histogram(
		aes(y=..density..),
		binwidth=0.2,
		fill="#0072B2",
		color="black"
	)
	+ stat_function(
		fun=dnorm,
		arg=list(
			mean=lambda^(-1),
			sd=(lambda*sqrt(exponentials))^-1),
			size=2)
	+ labs(
		title="Plot of the Simulations",
		x="Simulation Mean"
	)
```
