---
title: "Investigate the exponential distribution in R and compare it with the Central Limit Theorem"
author: "Glenn Kerbein"
date: "April 1, 2017"
output: pdf_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
The purpose of this analysis is to investigate the exponential distribution and compare it to the Central Limit Theorem. For this analysis, set ``lambda`` equal to ``0.2`` for all simulations. The investigation will compare the distribution of averages of 40 exponentials over 1000 simulations.

## Simulations

Set the simulation variables lambda, exponentials, and seed.

```{r echo=TRUE}
ECHO=TRUE
set.seed(9001)
lambda <- 0.2
exponentials <- 40
```

Run the simulations with variables.

```{r echo=TRUE}
simMeans = NULL
for (i in 1 : 1000) simMeans = c(simMeans, mean(rexp(exponentials, lambda)))
```

## Sample Mean versus Theoretical Mean

### Sample Mean


### Theoretical Mean


### Comparison


## Sample Variance versus Theoretical Variance

### Sample Variance

### Theoretical Variance

### Comparison

## Distribution

